---
title: "Marketing analytics A 1"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2023-02-26"
---
# Loading required libraries
```{r libraries}
library(readxl)
library(dplyr)
library(dlookr)
library(ggplot2)
library(openxlsx)
library(ggpubr)
library (cluster)
library(factoextra)

```

# setting Seed and uploading data

```{r Load Data}
set.seed(123)
seg <- read_excel("Restaurant Data.xlsx")
```

## Summary and Descriptive Statistics

```{r Descriptive Statistics}
summary(seg)
describe(seg)
DQ <- diagnose(seg)
DQ1 <- diagnose_numeric(seg)
DQ2 <- diagnose_outlier(seg)
```

## Dealing with DQ Issues and outliers

```{r Data Quality Issues}
# removing order freq and order value outliers

seg$avg_order_size[seg$avg_order_size>150] <- NA #replacing avg order size above 150 by NA
seg$avg_order_freq[seg$avg_order_freq>20] <- NA #replacing avg order freq above 20 by NA
seg <- na.omit(seg) #dropping rows with NA values

#Creating calculated field for Occupation

seg$Occupation <- rep(0, dim(seg)[1]) #creating a new column with 0 values
seg$Occupation [seg$Health ==1]<- 1 # replacing 0 with new values according to occupation code columns
seg$Occupation [seg$Finc ==1]<- 2
seg$Occupation [seg$Sales ==1]<- 3
seg$Occupation [seg$Advt ==1]<- 4 
seg$Occupation [seg$Edu ==1]<- 5 
seg$Occupation [seg$Cons ==1]<- 6 
seg$Occupation [seg$Eng ==1]<- 7 
seg$Occupation [seg$Tech ==1]<- 8
seg$Occupation [seg$Retail ==1]<- 9
seg$Occupation [seg$SMB ==1]<- 10

summary(seg)
dim(seg)
```
# Subsetting the bases variable for clustering
```{r Data subset for clustering}
new_seg <- seg[2:15]
   #choosing variable 2 to 15 (bases variables) from main data set
```



## CLUSTERING METHODS
#HIERARCHICAL CLUSTERING
# Hierarchical Clustering using Complete Linkage Method

```{r Hierarchical clustering- Complete Linkage}
set.seed(123)
hc.complete <- hclust(dist(scale(new_seg)), method = "complete") 

x1 <- c(1:10)
sort_height1 <- sort(hc.complete$height, decreasing=TRUE)
y1 <- sort_height1[1:10]
plot(x1,y1, xlab="number of clusters", ylab="Height", main= "Elbow plot using Complete linkage"); lines(x1,y1,col="blue")

#create a dendrogram of hierarchical cluster
fviz_dend(x=hc.complete, cex=0.7, lwd=0.7, k=4, k_colors = c("red","green","blue","magenta"), reg=T,
          rect_border ="grey",rect_fill=F )

hc.complete1 <- cutree(hc.complete,4) #cutting hierarchical cluster at 4th level from top bottom
table(hc.complete1) #shows number of observations in each cluster
```

# Hierarchical Clustering using Single Linkage Method
```{r Hierarchical clustering- Single Linkage}
set.seed(123)
hc.single <- hclust(dist(scale(new_seg)), method = "single")

x2 <- c(1:10)
sort_height2 <- sort(hc.single$height, decreasing=TRUE)
y2 <- sort_height2[1:10]
plot(x2,y2,xlab="number of clusters", ylab="Height", main= "Elbow plot using Single linkage"); lines(x2,y2,col="blue")

#create a dendrogram of hierarchical cluster
fviz_dend(x=hc.single, cex=0.7, lwd=0.7, k= 6, k_colors = c("red","green","blue","magenta","orange","purple"), reg=T,
          rect_border ="grey",rect_fill=F )

hc.single1 <- cutree(hc.single,6)
table(hc.single1)
```

# Hierarchical Clustering using Average Linkage Method
```{r Hierarchical clustering- Average Linkage}
set.seed(123)
hc.average <- hclust(dist(scale(new_seg)), method = "average")

x3 <- c(1:10)
sort_height3 <- sort(hc.average$height, decreasing=TRUE)
y3 <- sort_height3[1:10]
plot(x3,y3,xlab="number of clusters", ylab="Height", main= "Elbow plot using Average linkage"); lines(x3,y3,col="blue")

#create a dendrogram of hierarchical cluster
fviz_dend(x=hc.average, cex=0.7, lwd=0.7, k=3, k_colors = c("red","green","blue"), reg=T,
          rect_border ="grey",rect_fill=F )

hc.average1 <- cutree(hc.average,3)
table(hc.average1)
```

# Hierarchical Clustering using Centroid Linkage Method

```{r Hierarchical clustering- Centroid Linkage}
set.seed(123)
hc.centroid <- hclust(dist(scale(new_seg)), method = "centroid")

x4 <- c(1:10)
sort_height4 <- sort(hc.centroid$height, decreasing=TRUE)
y4 <- sort_height4[1:10]
plot(x4,y4,xlab="number of clusters", ylab="Height", main= "Elbow plot using Centroid linkage"); lines(x4,y4,col="blue")

#create a dendrogram of hierarchical cluster
fviz_dend(x=hc.centroid, cex=0.7, lwd=0.7, k=5, k_colors = c("red","green","blue","magenta","orange"), reg=T,
          rect_border ="grey",rect_fill=F )

hc.centroid1 <- cutree(hc.centroid,5)
table(hc.centroid1)
```


# Hierarchical Clustering using Ward Linkage Method
```{r Hierarchical clustering- Wards Linkage}
set.seed(123)
hc.ward <- hclust(dist(scale(new_seg)), method = "ward")

x5 <- c(1:10)
sort_height5 <- sort(hc.ward$height, decreasing=TRUE)
y5 <- sort_height5[1:10]
plot(x5,y5, xlab="clusters", ylab="height", main="Elbow plot using Ward's method"); lines(x5,y5,col="blue")

#create a dendrogram of hierarchical cluster
fviz_dend(x=hc.ward, cex=0.7, lwd=0.7, k=4, k_colors = c("red","green","blue","magenta"), reg=T,
          rect_border ="grey",rect_fill=F )

hc.ward1 <- cutree(hc.ward,4)
table(hc.ward1)
```
# elbow plot from above hierarchical clustering using different linkage methods mostly point out at 3, 4,5,6 clusters to be ideal number of clusters

  # K MEANS CLUSTERING
# for 3 clusters
```{r K means clustering 3 clusters }
set.seed(123)
new_seg1 <- scale(new_seg) #scaling data for standardisation

# k means algorithm for k =3
seg_kmeans1 <- kmeans(new_seg1,3, nstart=50)
seg_kmeans1

print("Centers")
seg_kmeans1$centers
print("total ss")
seg_kmeans1$totss
print("within ss")
seg_kmeans1$withinss
print("total within ss")
seg_kmeans1$tot.withinss
print("between ss")
seg_kmeans1$betweenss
print("size")
seg_kmeans1$size
print("iter")
seg_kmeans1$iter

segment_3c= seg_kmeans1$cluster #segregating cluster column from segmented dataset
seg1 <- seg
seg1 <- cbind(seg1,segment_3c) #binding new cluster column to original data set

a1 <- aggregate(new_seg, by=list(cluster=seg_kmeans1$cluster), mean) #calculating mean of clusters against all variables on unscaled data set

```

# for 4 clusters

```{r K means clustering 4 clusters }
set.seed(123)

#k means algorithm for k =4
seg_kmeans2 <- kmeans(new_seg1,4, nstart=50)
seg_kmeans2

print("Centers")
seg_kmeans2$centers
print("total ss")
seg_kmeans2$totss
print("within ss")
seg_kmeans2$withinss
print("total within ss")
seg_kmeans2$tot.withinss
print("between ss")
seg_kmeans2$betweenss
print("size")
seg_kmeans2$size
print("iter")
seg_kmeans2$iter

segment_4c= seg_kmeans2$cluster

seg1 <- cbind(seg1,segment_4c)

a2 <- aggregate(new_seg, by=list(cluster=seg_kmeans2$cluster), mean)

```

# for 5 clusters
```{r K means clustering 5 clusters }
set.seed(123)

#k means algorithm for k =5
seg_kmeans3 <- kmeans(new_seg1,5, nstart=50)
seg_kmeans3

print("Centers")
seg_kmeans3$centers
print("total ss")
seg_kmeans3$totss
print("within ss")
seg_kmeans3$withinss
print("total within ss")
seg_kmeans3$tot.withinss
print("between ss")
seg_kmeans3$betweenss
print("size")
seg_kmeans3$size
print("iter")
seg_kmeans3$iter

segment_5c= seg_kmeans3$cluster

seg1 <- cbind(seg1,segment_5c)

a3 <- aggregate(new_seg, by=list(cluster=seg_kmeans3$cluster), mean)

```

# for 6 clusters
```{r K means clustering 6 clusters }
set.seed(123)

# k means algorithm for k =6
seg_kmeans4 <- kmeans(new_seg1,6, nstart=50)
seg_kmeans4

print("Centers")
seg_kmeans4$centers
print("total ss")
seg_kmeans4$totss
print("within ss")
seg_kmeans4$withinss
print("total within ss")
seg_kmeans4$tot.withinss
print("between ss")
seg_kmeans4$betweenss
print("size")
seg_kmeans4$size
print("iter")
seg_kmeans4$iter

segment_6c= seg_kmeans4$cluster

seg1 <- cbind(seg1,segment_6c)

a4 <- aggregate(new_seg, by=list(cluster=seg_kmeans4$cluster), mean)
```


## VISUALISATIONS

# Silhouette plot & width  for 3 clusters
```{r}
sil1 <- silhouette(segment_3c, dist(new_seg1))
fviz_silhouette(sil1)
```

# Silhouette plot & width for 4 clusters
```{r}
sil2 <- silhouette(segment_4c, dist(new_seg1))
fviz_silhouette(sil2)
```

# Silhouette plot & width for 5 clusters
```{r}
sil3 <- silhouette(segment_5c, dist(new_seg1))
fviz_silhouette(sil3)
```

# Silhouette plot & width for 6 clusters
```{r}
sil4 <- silhouette(segment_6c, dist(new_seg1))
fviz_silhouette(sil4)
```


# Cluster Plots 
```{r}
# for 3 clusters
fviz_cluster(seg_kmeans1, new_seg1,
             palette = "Set2" , 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
)
# for 4 clusters
fviz_cluster(seg_kmeans2, new_seg1,
             palette = "Set3" , 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
)
# for 5 clusters
fviz_cluster(seg_kmeans3, new_seg1,
             palette = "Set4" , 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
)
# for 6 clusters
fviz_cluster(seg_kmeans4, new_seg1,
             palette = "Set5" , 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
)

```

#silhouette score plot against number of clusters

```{r} 
# creating a function to calculate silhouette score for clusters ranging from 2 to 10
silhouette_score <- function(k){set.seed(123)
  km <- kmeans(new_seg1, centers=k, nstart=50) #runs k means algorithm for every value of k iterating it 50 times
  ss <- silhouette(km$cluster, dist(new_seg1)) # calculating silhouette score for each observation
  mean(ss[,2]) # calculating mean of silhouette scores calculated above
}
k <- 2:5  # assigning number of clusters to k to be used in above function
avg_sil <- sapply(k,silhouette_score) #binding silhouette score against cluster numbers
avg_sil[2:4] # to display silhouette score of clusters 3,4 ,5 & 6
plot(k, type='b', avg_sil, xlab='Number of clusters', ylab='Average Silhouette Scores', frame=FALSE) # plotting average silhouette score against number of clusters.

```

#Linear Discriminant Analysis
```{r}
library(MASS)

#Discriminant Analysis 

lda.fit <- lda(segment_4c~ Education + Gender + Age + Income + zip_code + Occupation, data = seg1)
lda.fit

#Calculating the discriminant score - Predicting the segments based on discriminant score

lda.pred <- predict(lda.fit, seg1)
names(lda.pred)
seg1$lda.class<-lda.pred$class
tclass<-table(seg1$segment_3c, seg1$lda.class)
tclass
sum(diag(tclass))/nrow(seg1)#accuracy of the confusion matrix
```


```{r}
set.seed(123)
fviz_nbclust (new_seg1, kmeans, method = "wss") + geom_vline(xintercept = 4, linetype = 2)
fviz_nbclust (new_seg1, kmeans, method = "silhouette") + geom_vline(xintercept = 4, linetype = 2)
fviz_nbclust (new_seg1, kmeans, method = "gap_stat") + geom_vline(xintercept = 4, linetype = 2)
```

# write excel back to system
```{r}
segnew <- seg[-(16:25)]  #dropping columns not required
segnew <- cbind(segnew,segment_4c)

write.xlsx(segnew, '/Users/jyotismac/Documents/BUSINESS ANALYTICS/DATA SETS')
write.xlsx(a2, '/Users/jyotismac/Documents/BUSINESS ANALYTICS/DATA SETS')


```
# to rknit the markdown file
```{r}

```

